Investigation of the Effect of Pretraining on EEG Classification with Deep Recurrent-Convolutional Networks

### 2017-10-15 paper/paper.tex
Fix the last messed up merge

For our classification we used several layers of LSTMs, because the temporal dependencies of our data are little understood. Unidirectional and bidirectional LSTMs were tried, the latter ones learn much faster and more. Bidirectionality was implemented by simply merging the final hidden states of a normal LSTM with an LSTM with inputs reversed. Several number of layers and dropout rates were tried, $2$ layers and a $0.1$ dropout scored highest. These relatively small numbers could both be due to the small dataset. A little improvement was gained by replacing hard-sigmoid activation for the input with sigmoid activation. All models were trained for $50$ epochs over the whole training set. The models were first trained without the weights of the decoder and then compared to the model with trained decoder.

@@

 SVM & - & 7.78\% $\pm 4.1\% $\\ 
 Random Forest & - & 36.72\% $\pm 7.3\%$ \\ 
 Logistic Regression & - & 33.6\% $\pm 5.8\%$ \\ 

@@

 2 layer unidirectional LSTM & - & 18.55\% $\pm \% $\\ 
 2 layer bidirectional LSTM & - & 83.68\% $\pm \%$ \\ 
 3 layer bidirectional LSTM & - & 83.39\% $\pm \%$ \\ 
 5 layer bidirectional LSTM & - & 75.78\% $\pm \%$ \\
 2 layer bidirectional LSTM with 0.1 dropout & - & 84.08\% $\pm \%$ \\
 2 layer bidirectional LSTM with 0.2 dropout & - & 79.87\% $\pm \%$ \\

@@

 2 layer bidirectional LSTM with 0.2 dropout & - & 89.21\% $\pm \%$ \\

@@

We could also gain some benefit by doing introspection, or analyzing the internal state of our trained model to see what representations it has actually learned. Bashivan et al. use a deconvolution method to look at feature maps and back projections for inputs. Using this method and working with neuroscientists familiar with brain activity patterns, we could determine the significance of what is being learned in the model and possibly use this improve our model, ie find hyperparameters that remove unrelated mental activity.
\subsection{Future investigations}
There are many directions for future work uncovered during the implementation of our project. One simple idea we could easily investigate with our model is performance on classification with different labels. Since the dataset already contains labels for mental tasks, we could easily evaluate our network to classify EEG recordings on these tasks. There is also a high amount of data available in the dataset that does not correspond to class labels, for example dream descriptions, which are simple text reports. It may be an interesting idea to investigate modifying the network to make use of this data, such as generation of text for an EEG input sequence. 
\begin{thebibliography}{99}

@@

    © 2017 GitHub, Inc.
    Terms
    Privacy
    Security
    Status
    Help

    Contact GitHub
    API
    Training
    Shop
    Blog
    About

>>>>

For our classification we used several layers of LSTMs, because the temporal dependencies of our data are little understood. We wanted to allow the model to classify according to anomalies or structure of the sequence. Unidirectional and bidirectional LSTMs were tried, the latter ones learn much faster and more. This is probably due to uncertainty where in the sequence the rich part of the data is, unlikelihood that is at similar positions and the fact that the data is not temporally resolving but just developing so it can be viewed equally from both directions. Several number of layers and dropout rates were tried, $2$ layers and a $0.1$ dropout scored highest. These relatively small numbers could both be due to the small dataset. A little improvement was gained by replacing hard-sigmoid activation for the input with sigmoid activation. All models were trained for $50$ epochs over the whole training set. The models were first trained without the weights of the decoder and then compared to the model with trained encoder.

@@

 SVM & - & 7.8\% $\pm \, 4.1\% $\\ 
 Random Forest & - & 36.7\% $\pm \, 7.3\%$ \\ 
 Logistic Regression & - & 33.6\% $\pm \, 5.8\%$ \\ 

@@

 2 layer unidirectional LSTM & - & 18.6\% $\pm \, 2.0\% $\\ 
 2 layer bidirectional LSTM & - & 83.7\% $\pm \, 4.1\%$ \\ 
 3 layer bidirectional LSTM & - & 83.4\% $\pm \, 3.0\%$ \\ 
 5 layer bidirectional LSTM & - & 75.8\% $\pm \, 5.6\%$ \\
 2 layer bidirectional LSTM with 0.1 dropout & - & 84.1\% $\pm \, 4.2\%$ \\
 2 layer bidirectional LSTM with 0.2 dropout & - & 79.9\% $\pm \, 5.0\%$ \\

@@

 2 layer bidirectional LSTM with 0.1 dropout & - & 89.2\% $\pm \, 4.4\%$ \\

@@

We could also gain some benefit by doing introspection, or analyzing the internal state of our trained model to see what representations it has actually learned. Bashivan et al. use a deconvolution method to look at feature maps and back projections for inputs. Using this method and working with neuroscientists familiar with brain activity patterns, we could determine the significance of what is being learned in the model and possibly use this improve our model, ie find hyperparameters that remove unrelated mental activity.
\subsection{Future investigations}
There are many directions for future work uncovered during the implementation of our project. One simple idea we could easily investigate with our model is performance on classification with different labels. Since the dataset already contains labels for mental tasks, we can easily evaluate our network to classify EEG recordings on these tasks. Some preliminary results have been found for this classification task: employing a subject-out cross-validation evaluation scheme, we have achieved an average accuracy of $51.31\%$, which, given that the distribution of the 3 different tasks was approximately $50\%$, $25\%$, $25\%$, is not significantly higher than a random "majority class" guess. This would indicate that the per-subject variations are much easier to pick up for the network than those related to a specific task, however, further investigation into this area would probably greatly improve results.

There is also a high amount of data available in the dataset that does not correspond to class labels, for example dream descriptions, which are simple text reports. It may be an interesting idea to investigate modifying the network to make use of this data, such as generation of text for an EEG input sequence. 
\begin{thebibliography}{99}

@@
